---
title: "Code Portfolio: Week 5 to 7"
author: "Katherin Jardim Rodrigues"
date: "December, 2018"
output: 
  html_document:
    toc: TRUE # table of content true
    toc_float: TRUE #table of content in a side
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, warning=FALSE)
```

# Week 5

##Week.5: Files load

*Packages:* readr and tidyverse for flat files 

**Remember to load a package, you should use the function 'library()'**

Load flat files in R with the readr package:   
library(tidyverse)  
read.csv() *comma delimited files*  
read.csv2() *reads semicolon separated files*   
read_tsv() *tabs delimited files*  
read.delim() *reads in files with any delimiter*  
read_fwf() *reads fixed width files*  

Every single file type that we want to get into R, require its own function.

**Recommendations:**  
- If we are going to work with Excel, the first row is usually reserved for the header, and the first column for the sampling unit.  
- We cannot use blank spaces in the headers. Below, we are going to mention some functions to avoid this.  
- Short names are preferred over longer names. 
- Avoid special characters.  
- Remember to delete the comments (we can use 'skip=')  
- Verify that missing values are indicated as NA.   
- To start a work, we can use 'rm(list=ls())' to remove data and values previously processed. 

```{r eval=FALSE}
#"\n" -> Shortcut for adding a new line
a -> read_csv("1,2,3\n4,5,6", col_names = FALSE)
```

**What function would you use to read a file where fields were separated with "|"?**
```{r eval=FALSE}
read_delim(file, delim = "|")
```

**Apart from file, skip, and comment, what other arguments do read_csv() and read_tsv() have in common?**

-   `col_names` and `col_types`
-   `locale` determine if "." or "," is used as a decimal mark.
-   `na` and `quoted_na` missing values
-   `trim_ws` trims whitespace 
-   `n_max` rows to read
-   `guess_max` Maximum number of records to use for guessing column types
-   `progress` Display a progress bar? By default it will only display in an interactive session and not while knitting a document. The display is updated every 50,000 values and will only display if estimated reading time is 5 seconds or more. The automatic progress bar can be disabled by setting option readr.show_progress to FALSE.

Reference -> https://readr.tidyverse.org/reference/read_delim.html

**What are the most important arguments to read_fwf()?**  
col_positions	
Column positions, as created by fwf_empty(), fwf_widths() or fwf_positions().

Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like " or '. By convention, read_csv() assumes that the quoting character will be ", and if you want to change it you'll need to use read_delim() instead.  

**What arguments do you need to specify to read the following text into a data frame?**
```
"x,y\n1,'a,b'"
```
`x <- "x,y\n1,'a,b'"
read_csv(x, quote = "'")`

**Identify what is wrong with each of the following inline CSV files. What happens when you run the code?**

read_csv("a,b\n1,2,3\n4,5,6") #The first row is missing a value  
read_csv("a,b,c\n1,2\n1,2,3,4") #The third column is missing a value  
read_csv("a,b\n\"1") #The second column is missing a value  
read_csv("a,b\n1,2\na,b") #create character table  
read_csv("a;b\n1;3") #the values have semicolons instead of comma  

**What are the most important arguments to locale()?**  
-   date and time formats: `date_names`, `date_format`, and `time_format`  
-   time zone: `tz`  
-   numbers: `decimal_mark`, `grouping_mark`  
-   encoding: `encoding`  

**What happens if you try and set decimal_mark and grouping_mark to the same character? What happens to the default value of grouping_mark when you set decimal_mark to ","? What happens to the default value of decimal_mark when you set the grouping_mark to "."?**  
1. It's going to generate an error.  
2. The grouping mark will be period `"."`    
3. The decimal mark will be a comma  

**Other notes**  

- We can rewrite the names of the columns to remove any spaces. 
'names(ozone) <- make.names(names(ozone))'
- 'str' -> number of rows and columns, classes. 
- Look at the Top and the Bottom of your data: 'head()' and 'tail()'  

##Week.5: Practice

```{r}
library(readr)
ipeds2016 <- read_csv("C:/Users/Katherin/Desktop/MSC ANALYTICS/ANLY506/CodePortfolio/Data/ipeds2016.csv") #Import the dataset .csv with readr and establish the first row as column names
#str(ipeds2016)  #Verify classes, number of rows and columns

```

```{r}
head(ipeds2016)  #First rows to verify format and information
#tail(ipeds2016)  #Last rows to verify format and information
```

# Week 6  

Reference: https://bookdown.org/ndphillips/YaRrr/matricesdataframes.html 

## Week.6: Creating matrices and dataframes  
- cbind(a, b, c) *combine vectors as columns in a matrix* 
```{r}
cbind(1:5, 6:10, 11:15)
```
- rbind(a, b, c) *combine vectors as rows in a matrix*
```{r}
rbind(1:5, 6:10, 11:15)
```
-matrix(x, nrow, ncol, byrow) *create a matrix from a vector 'x'
```{r}
x <- 1:12
matrix(x, nrow = 3, ncol = 4)
```
- data.frame() *create a dataframe from named columns* 
```{r}
data.frame("age" = c(19,21), "sex" =c("m", "f"), stringsAsFactors = FALSE)
#In dataframe, we should add stringsAsFactors=FALSE to maintain sex as character
```
We can access specific data using brackets: 'data[rows, columns]

##Week.6: Practice

- *Combine the data into a single dataframe*
```{r}
pirates <- data.frame("Name" = c("Astrid", "Lea", "Sarina", "Ramon", "Letizia", "Babice", "Jonas", "Wendy", "Niveditha", "Gioia"), "Sex" = c("F", "F","F","M","F","F","M","F","F","F"), "Age" = c(30, 25, 25, 29, 22, 22, 35, 19, 32, 21), "Superhero" = c("Batman", "Superman", "Batman", "Spiderman", "Batman", "Antman", "Batman", "Superman", "Magott", "Superman"), "Tattoos"= c(11, 15, 12, 5, 65, 3, 9, 13, 900, 0), stringsAsFactors = FALSE)
head(pirates) 
 
```

- *What is the median age of the 10 pirates?*
```{r}
nrow(pirates) #to validate that they are 10
median(pirates$Age) #Using $ to calculate the median
with(pirates, median(Age)) #Using function 'With()'
```
-*What was the mean age of female and male pirates separately?*
```{r}
Females <- mean(pirates$Age[pirates$Sex == "F"])
Males <- mean(pirates$Age[pirates$Sex =="M"])
Mean.Age <- cbind(Females, Males)
print(Mean.Age)
```
-*What was the most number of tattoos owned by a male pirate?*
```{r}
max(pirates$Tattoos[pirates$Sex=="M"])
```
-*What percent of pirates under the age of 32 were female?*
```{r}
with(pirates, mean(Sex[Age < 32] == "F"))
```

-*What percent of female pirates are under the age of 32?*
```{r}
with(pirates, mean(Age[Sex=="F"] < 32))
```


-*Add a new column to the dataframe called tattoos.per.year which shows how many tattoos each pirate has for each year in their life.*
```{r}
pirates$tattoos.per.year <- with(pirates, Tattoos/Age)
tail(pirates)
```

-*Which pirate had the most number of tattoos per year?*
```{r}
pirates$Name[pirates$tattoos.per.year==max(pirates$tattoos.per.year)]
```

-*What are the names of the female pirates whose favorite superhero is Superman?*
```{r}
pirates$Name[pirates$Superhero=="Superman"]
```

-*What was the median number of tattoos of pirates over the age of 20 whose favorite superhero is Spiderman?*
```{r}
median(pirates$Tattoos[pirates$Age>20 & pirates$Superhero=="Spiderman"])
```

# Week 7

Reference: https://r4ds.had.co.nz/transform.html

## Week.7: Data transformation  

*Packages:* dplyr 

- **Logical vectors** are: `&` -> and; `|` -> or; `==` -> equal; `!` -> not.  
-  To determine if a value is missing, use `is.na()`  

##Week.7: Practice

* *Find all flights that*

- *Had an arrival delay of two or more hours*
```{r}
library(dplyr)
flights <- nycflights13::flights 
arrival.delay <- filter(flights, arr_delay>=120)
head(arrival.delay)
```

- *Flew to Houston (IAH or HOU)*
```{r}
houston <- filter(flights, dest == "IAH" | dest == "HOU")
tail(houston)
```

- *Were operated by United, American, or Delta* 
```{r}
airlines <- filter(flights, carrier %in% c('UA', 'AA', 'DL'))
head(airlines)
```

- *Departed in summer (July, August, and September)*
```{r}
summer <- filter(flights, month %in% c(7, 8, 9))
head(summer)
```

- *Arrived more than two hours late, but didn't leave late*
```{r}
late <- filter(flights, arr_delay > 120, dep_delay <= 0)
head(late)
```

- *Were delayed by at least an hour, but made up over 30 minutes in flight*
```{r}
delayed <- filter(flights, dep_delay >= 60, dep_delay-arr_delay > 30)
tail(delayed)
```

-*Departed between midnight and 6am (inclusive)*
```{r}
departure <- filter(flights, dep_time <= 600 | dep_time == 2400)
head(departure)
```

-*Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?*

This is a shortcut for x >= left & x <= right, implemented efficiently in C++ for local values, and translated to the appropriate SQL for remote tables.

-*How many flights have a missing dep_time? What other variables are missing? What might these rows represent?*
8255 flights have missing dep_time. Other variables are: dep_delay, arr_time, arr_delay, air_time. These rows might represent canceled flights, rerouting or a normal lost data. 
```{r}
summary(flights)
```

-**Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)**
 NA ^ 0 -> equals to 1  
 NA | TRUE -> the `|` operator returns TRUE if either of the terms are true  
 FALSE & NA -> `&` returns TRUE when both terms are true.
 