---
title: "Code Portfolio: Week 10 to 12"
author: "Katherin Jardim Rodrigues"
date: "February, 2019"
output: 
  html_document:
    toc: TRUE # table of content true
    toc_float: TRUE #table of content in a side
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, warning=FALSE)
```

# Week 10

##Week.10: Other visualization tools & Data summary  
- We can adjust the size of margins by specifying a margin parameter using `par(mar)` -> par(mar = c(top, right, bottom, left)).  
- If we want to do one dimension summaries, we can use: five number summary, histogram and boxplot.  
- If we want to do with two dimensions: Multiple/overlayed (lattice/ggplot2), scatterplots, smooth scatterplots. 

##Week.10: Practice 
```{r}
library(ggplot2)
par(mfrow = c(1, 3), mar = c(4, 4, 2, 1), oma = c(0, 0, 2, 0))
with(diamonds, {
 plot(carat, price, main = "Carat and Price")
 plot(depth, price, main = "Price and depth")
 plot(table, price, main = "Price and Table")
 mtext("Price and its features in Diamonds", outer = TRUE)
 })

```


# Week 12

Reference: https://uc-r.github.io/kmeans_clustering

##Week.12: K-means Cluster Analysis

*Packages:* tidyverse / cluster & factoextra  

- Rows are observations (individuals) and columns are variables  
- Any missing value in the data must be removed or estimated.  
- The data must be standardized (i.e., scaled) to make variables comparable. Recall that, standardization consists of transforming the variables such that they have mean zero and standard deviation one  
- To remove any missing value that might be present in the data, type this: `df <- na.omit(df)`

```{r}
df <- USArrests
df <- na.omit(df)
df <- scale(df)
head(df)
```

```{r error=FALSE, warning=FALSE}
library(factoextra)
distance <- get_dist(df)
fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))
```



```{r}
library(cluster)
# compute gap statistic
set.seed(123)
gap_stat <- clusGap(df, FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
# Print the result
print(gap_stat, method = "firstmax")
fviz_gap_stat(gap_stat)
```

```{r}
set.seed(123)
final <- kmeans(df, 4, nstart = 25)
print(final)
fviz_cluster(final, data = df)
```